\documentclass[11pt]{article}
\usepackage{authblk}
\input{preamble}
\input{values}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,      
    urlcolor=blue,
    citecolor=blue,
}

\usepackage{xcolor}
\newcommand{\bpn}[1]{\textcolor{blue}{\textbf{[BPN: #1]}}}
\newcommand{\jrb}[1]{\textcolor{violet}{\textbf{[JRB: #1]}}}


%\title{Unbinned OmniFold measurement of the MINERvA ME FHC\\
%CC-inclusive double-differential cross section\\
%$d^{2}\sigma/(d\pT\,d\pPar)$ and its higher-dimensional extensions\\[4pt]
%\large MINERvA analysis note --- DRAFT}

\title{\textbf{Analysis Note:}\\
Unbinned measurement of the charged-current inclusive cross section in vN in collisions with public MINERvA data}

\author[1]{Joseph Bailey\thanks{\url{josephrb@stanford.edu}}}
\author[2]{Benjamin Nachman\thanks{\url{nachman@stanford.edu}}}
\affil[1]{Department of Physics, Stanford University}
\affil[2]{Department of Particle Physics and Astrophysics, Stanford University and SLAC National Laboratory}


\date{\today\\[4pt]\small Code and file references: \repolink}

\begin{document}
\maketitle

\begin{abstract}
\bpn{The abstract is too long - let's put all the detailed content in the executive summary and then keep the abstract shorter and more high level.}
\jrb{Took out numbers and just introduced each section.}

We apply \emph{unbinned} OmniFold to the MINERvA medium-energy FHC charged-current inclusive data. We first reproduce the published double-differential cross section $d^{2}\sigma/(d\pT\,d\pPar)$~\cite{MINERvA:2021owq}, recovering both the spectrum and its uncertainty budget with a fully independent method. We then perform the first simultaneous three-, four-, and five-observable unfolds of MINERvA data, adding available energy ($\Eavail$), three-momentum transfer ($q_{3}$), and hadronic invariant mass ($W$); each reproduces its lower-dimensional predecessor under marginalization. These higher-dimension results recover the established MINERvA low-recoil/2p2h deficit through an independent method and localise a residual high-$\Eavail$ excess to the deep-inelastic region, where no current generator reproduces the data. We further demonstrate a point-cloud transformer that unfolds raw reconstructed calorimeter clusters into an absolutely-normalized cross section, and a first full-phase-space CC-inclusive measurement obtained by removing the muon kinematic cuts. Quantitative results, the full uncertainty budgets, and the validation suite are given in the executive summary and the sections that follow. 

\end{abstract}

\input{sec_execsummary}
\clearpage

\tableofcontents

\input{sec_intro}
\input{sec_experiment}
\input{sec_method}
\input{sec_results}
\input{sec_systematics}
\input{sec_validation}
\input{sec_3d}
\input{sec_eavailw}
\input{sec_pet}
\input{sec_fps}
\input{sec_summary}

\section*{Acknowledgments}
I thank my advisor, Benjamin Nachman, for guidance throughout this work. This
analysis was carried out on the Perlmutter system at the National Energy
Research Scientific Computing Center (NERSC), a U.S.\ Department of Energy
Office of Science User Facility, under project m3246. It builds on the MINERvA
Analysis Toolkit (MAT) and the MINERvA-101 cross-section framework, and I thank
their authors. I am grateful to the MINERvA Collaboration for the public release
of the medium-energy CC-inclusive data product on which the reproduction rests.

\appendix
\input{app_statmethods}
\input{app_landscape}

\printbibliography

\end{document}
