\section{Method and analysis pipeline}
\label{sec:method}

\paragraph{Unbinned OmniFold.}
OmniFold~\cite{Andreassen:2019cjw,Andreassen:2021zzk} alternates two reweighting
steps per iteration: step~1 reweights the
reconstructed simulation to the data in observable space (a learned
data/MC likelihood ratio, pulled back to truth through the detector response),
and step~2 reweights truth-level simulation to the step-1 result, yielding a
self-consistent truth-level weight per event. We use gradient-boosted decision
trees as the classifiers/regressors. The engine is
\texttt{unbinned\_unfolding/python/omnifold.py}; the MINERvA driver is
\texttt{2d-unfolding/unfold\_2d\_omnifold\_unbinned.py} with backends
\texttt{\{exact,hist,lgbm\}} (\S\ref{sec:results} compares them). The unfolded
observables are $(\pT,\pPar)$ of the muon; the phase space is $\theta_{\mu}<20^{\circ}$,
$\pT<\SI{4.5}{GeV/c}$, $1.5<\pPar<\SI{60}{GeV/c}$.

\paragraph{Production pipeline (Phase~18.2).}
The C++ event loop on the MINERvA detector~\cite{MINERvA:2013zvz} with the NuMI
flux of Ref.~\cite{MINERvA:2016iqn}
(\texttt{MINERvA101/MINERvA-101-Cross-Section/runEventLoopOmniFold.cpp}) writes
the per-event truth and reconstructed trees consumed by the driver. Two
invariants make the result truth-tree authoritative:
\begin{itemize}
  \item \textbf{Truth-tree-authoritative reco gate}: a reconstructed signal event
        is filled only when its $(\mathrm{run},\mathrm{subrun},\mathrm{event})$
        key is present in the truth denominator; truth-only events enter as native
        OmniFold ``miss'' entries.
  \item \textbf{Completeness} $c = (\text{events in the OmniFold input})/(\text{all
        truth events})$ is $1.000000$ by construction at ME~FHC
        (\num{32849103} entries), so the cross-section completeness division is a
        no-op self-check.
\end{itemize}

\paragraph{Cross-section extraction.}
Unfolded truth-level counts are converted to the differential cross section via
$d^{2}\sigma = U/(c\,\Phi\,N\,\mathrm{POT}\,\Delta\pT\,\Delta\pPar)\times10^{4}$,
where $\Phi$ is the per-$\pT$ flux integral (broadcast over $\pPar$), $N$ the
fiducial nucleon count, and POT the data exposure
(\texttt{extract\_cross\_section\_2d}). The flux normalisation enters as
$\sigma\propto1/\Phi$; correctly varying $\Phi$ per flux universe was the
dominant systematic fix (\S\ref{sec:systematics}).

\paragraph{Binning.}
The reported binning (14 $\pT$ $\times$ 16 $\pPar$ bins; 205 of 224 reported)
follows \texttt{minerva\_paper\_anc/bin\_mapping.txt}; the 19 unreported diagonal
cells have $\pT^{\mathrm{hi}}/\pPar^{\mathrm{lo}}>\tan20^{\circ}$.
\TODO{insert the binning table or cite uq\_statistical\_methods.tex.}
