\section{Introduction and motivation}
\label{sec:intro}

The MINERvA medium-energy (ME) FHC charged-current (CC) inclusive
double-differential cross section in muon transverse and longitudinal momentum,
$d^{2}\sigma/(d\pT\,d\pPar)$, was published by Ruterbories \emph{et al.}
\cite{MINERvA:2021owq} using D'Agostini iterative Bayesian unfolding
(IBU)~\cite{DAgostini:1994fjx} on a binned response matrix. This note reproduces
that measurement with \emph{unbinned} OmniFold~\cite{Andreassen:2019cjw,
Andreassen:2021zzk}, an unfolding technique based on iterated reweighting with
machine-learned likelihood ratios that operates on per-event observables rather
than on a fixed histogram binning. First demonstrated experimentally by the H1
collaboration~\cite{H1:2021wkz,H1:2023fzk}, OmniFold is one of a family of
machine-learning unfolding methods~\cite{Datta:2018mwd,Bellagente:2019uyp,
Bellagente:2020piv,Backes:2022vmn,Chan:2023tbf} that lift the binned-response
limitations of classical IBU, the SVD method~\cite{Hocker:1995kb} and
TUnfold~\cite{Schmitt:2012kp} (the latter accessed here via
RooUnfold~\cite{Adye:2011gm}).

\paragraph{Why unbinned OmniFold.}
\begin{itemize}
  \item \textbf{No binning at unfold time.} OmniFold learns event weights in the
        full observable space; the result is histogrammed only at the end, so the
        binning can be chosen (or refined) after unfolding without re-running.
  \item \textbf{Natural path to higher dimensions.} Adding observables is adding
        input features, not multiplying response-matrix bins --- this is the
        enabling property for the available-energy extension
        (\S\ref{sec:3d}) that has no practical IBU analogue.
  \item \textbf{Different regularisation.} OmniFold's regulariser is the
        iteration count and the model capacity of the gradient-boosted decision
        trees (GBDTs), not the IBU prior-iteration coupling. Comparing the two
        on the same data quantifies the method-dependence of the result
        (\S\ref{sec:results}).
\end{itemize}

\paragraph{Goal and scope.}
The goal is a faithful, fully uncertainty-quantified reproduction of the
published 2D cross section as the demonstrator for unbinned OmniFold on MINERvA
data, followed by its higher-dimensional extensions. This note covers the
experiment and samples (\S\ref{sec:experiment}), the method and pipeline
(\S\ref{sec:method}), the 2D result and its comparison to the published data
and to GENIE MINERvA Tune v1 (\S\ref{sec:results}), the uncertainty budget
(\S\ref{sec:systematics}), the validation suite (\S\ref{sec:validation}), the
three- and four-observable extensions with available energy and $q_{3}$
(\S\ref{sec:3d}), the $(\Eavail,W)$ study that localises the remaining
deep-inelastic-tail excess (\S\ref{sec:eavailw}), the point-cloud transformer
track (\S\ref{sec:pet}), and a summary with the planned full-phase-space
extension (\S\ref{sec:summary}). Appendix~\ref{sec:openquestions} records the
resolution of the questions raised during internal review;
Appendix~\ref{app:statmethods} is the complete statistical methodology.
