\section{Resolution of the open questions from internal review}
\label{sec:openquestions}
\label{app:openq}

An earlier draft listed six items for collaboration input. The 2026-06-03 audit
(scripts under \texttt{2d-unfolding/uq/} and \texttt{3d-unfolding/genie/}; full
log in \texttt{LITERATURE\_NOTES.md}) closes or reduces all of them. Each item
below states its resolution and what, if anything, still needs a collaborator.
Technical detail is in App.~\ref{sec:rank}.

\begin{enumerate}
  \item \textbf{Flux\,$\leftrightarrow$\,Muon\_Energy joint block and the rank
        gap --- \emph{resolved}.} The premise of the original question (``the
        paper sums more bands than we propagate'') was wrong. Our systematic-only
        covariance is rank $140/205$; the paper's
        ($\texttt{TotalCov}-\texttt{StatOnly}$) is rank $\mathbf{120}$ ---
        \emph{lower} --- so we propagate at least as many independent systematic
        directions, using the same \texttt{GetStandardSystematics}. The paper's
        $V$ is full-rank only because it includes the \emph{statistical} block; we
        match that with $\mathbf{C}^{\mathrm{syst}}+\mathbf{C}^{\mathrm{boot}}$
        (rank $201$). The one named cross-band correlation, the Bashyal
        flux\,$\leftrightarrow$\,muon-energy-scale block~\cite{Bashyal:2021tqu},
        \emph{is} rederivable and now built
        (\texttt{uq/rederive\_flux\_muonE\_cross.py}\,$\to$\,\texttt{uq/flux\_muonE\_cross.root}):
        the muon-energy-scale systematic is rank-1, so the cross term is rank-2,
        with $\rho=0.900$ (two independent estimators, $0.87$--$0.90$). It is
        physically real but adding it leaves the rank unchanged ($140\to140$) ---
        its directions already lie in $\mathrm{range}(\mathbf{C}^{\mathrm{syst}})$.
        The residual stat-block difference (our bootstrap $2.5\times$ smaller by
        trace) is explained by the data/MC split bootstrap: it is present in the
        \emph{data-only} stream ($\mathbf{C}_{\mathrm{data}}/\texttt{StatOnly}=0.356$),
        i.e.\ genuine unbinned-OmniFold statistical efficiency, not a missing MC
        term. \emph{Remaining (optional):} the fitted
        $(\Phi_{\mathrm{norm}},\Phi_{\mathrm{shape}},\mathrm{MuonE}_{\mathrm{scale}})$
        posterior from MAT-MINERvA would give a fully paper-independent $\rho$.

  \item \textbf{FrInel\_pi exclusion --- \emph{resolved (negligible)}.} An
        independent \texttt{grwght1p} pass on the 2M CV events shows both pion-FSI
        dials are sub-percent on $d\sigma/d\Eavail$: \texttt{FrInel\_pi}
        $\le\SI{0.74}{\percent}$ and \texttt{FrAbs\_pi} $\le\SI{0.82}{\percent}$
        (\S\ref{sec:3d-2p2h}). A $<\SI{1}{\percent}$ knob cannot bridge the
        $7$--$\SI{15}{\percent}$ low-$\Eavail$ excess, and we are consistent with
        MAT in excluding it. \emph{Remaining (immaterial):} a one-line collaborator
        confirmation that the MAT exclusion is still current guidance.

  \item \textbf{PPFX effective rank --- \emph{resolved}.} The $100$-universe flux
        band covariance has hard rank $99$ but a participation-ratio effective rank
        of $\approx1$: a single coherent normalization mode carries
        \SI{99.6}{\percent} of the variance ($9$ modes for \SI{99.9}{\percent}).
        So the PPFX flux uncertainty is, to excellent approximation, a fixed-rank
        (rank-1, at most rank-9) object; treating the $100$ universes as $100$
        independent directions over-counts the flux dimensionality. This is also
        why the leading flux mode correlates $\rho\approx0.9$ with the rank-1
        muon-energy-scale mode (item~1). A PCA / fixed-rank flux covariance is
        therefore both justified and recommended.

  \item \textbf{Ours-only $\chi^{2}$ --- \emph{method resolved; precedent open}.}
        The obstruction is statistical, not rank: $\mathbf{C}^{\mathrm{ours}}=
        \mathbf{C}^{\mathrm{syst}}+\mathbf{C}^{\mathrm{boot}}+\mathbf{C}^{\mathrm{ML}}$
        is rank $201$ and invertible; we report the ours-only goodness-of-fit on it
        via a truncated-spectral $\chi^{2}$ (retain
        $\lambda>10^{-10}\lambda_{\max}$), the same convention as the 3D comparison,
        and lead with per-bin pulls plus the regularization scan
        (\S\ref{sec:results-tension}). \emph{Remaining:} published precedent from
        other MAT-based OmniFold/unbinned analyses for the ours-only $\chi^{2}$
        convention.

  \item \textbf{3D systematic covariance --- \emph{method resolved; precedent
        open}.} The available-energy extension carries a full combined covariance
        (universe$+$bootstrap$+$ML) over all $1431$ reported bins, rank $247$
        (\S\ref{sec:3d-syst}); the goodness-of-fit is the rank-$247$
        truncated-spectral $\chi^{2}$ (\S\ref{sec:3d-models}), the recommended
        treatment for a rank-deficient unbinned covariance. \emph{Remaining:}
        whether there is MINERvA precedent for \emph{publishing} a 3D systematic
        covariance from low-recoil samples, and endorsement of this GoF.

  \item \textbf{High-$\Eavail$ DIS-tail excess --- \emph{localized in $W$,
        generator- and tune-robust; modelling open}.} (This study is presented
        fully in \S\ref{sec:eavailw}; the audit record is kept here.)
        Enabling Valencia 2p2h resolves the
        low-recoil excess but leaves a separate $+2.2\sigma$ data excess in the
        $[1.50,3.00)\,\si{GeV}$ bin (\S\ref{sec:3d-2p2h}), where 2p2h does not
        contribute and the (sub-percent) FSI dials of item~2 cannot reach. To test
        the deep-inelastic hypothesis directly we added the hadronic invariant mass
        $W$ as a validated fifth OmniFold axis (its $W$-marginal recovers the frozen
        4D cross section to \SI{0.11}{\percent}) and compared the unfolded
        $(\Eavail,W)$ cross section to the GENIE central value. The excess is
        \emph{predominantly deep-inelastic}: the two highest-$\Eavail$ bands carry
        \SI{57}{\percent} of the positive excess; of all positive excess,
        high-$\Eavail$ ($\ge\SI{0.8}{GeV}$) holds \SI{67}{\percent} and
        \SI{83}{\percent} of \emph{that} sits at $W\ge\SI{1.8}{GeV}$, with the single
        largest cell the deep-DIS corner ($\Eavail>\SI{3}{GeV}$, $W>\SI{3}{GeV}$) at
        \SI{22}{\percent}. A secondary low-$W$ ($<\SI{1.1}{GeV}$, QE-like) excess and
        a small $\Delta$-region ($W\simeq1.4$--$\SI{1.8}{GeV}$) deficit are also
        present. So the missing strength behaves like a deep-inelastic-tail
        \emph{modelling} deficit, exactly the hypothesis the $W$ axis was added to
        test. We confirm this against a \emph{generator band}: independent
        spline-normalised $(\Eavail,W)$ predictions from GENIE~CV,
        GENIE\,$+$\,Valencia~2p2h/MEC, NuWro, and GiBUU all underpredict the
        high-$\Eavail\times$high-$W$ corner by \SIrange{54}{58}{\percent}
        (data/generator $=1.54,\,1.58,\,1.56$). Crucially, enabling Valencia 2p2h
        does \emph{not} close it --- it slightly \emph{worsens} the corner
        ($1.54\to1.58$), since 2p2h populates low $W$ --- and the independent NuWro
        and GiBUU generators miss it by the same or a larger margin (GiBUU is the
        most deficient overall, $\sigma=\SI{2.22e-38}{cm^2/nucleon}$); at
        $W\in[2.2,3.0)\,\si{GeV}$ all four sit \SIrange{23}{25}{\percent} below the
        data. The DIS-tail excess is therefore \emph{generator- and tune-robust}: no
        current model reproduces it. Attaching a number, we marginalise the adopted
        combined 4D covariance (\S\ref{sec:3d-syst}, unified-throw systematic) to the
        $\Eavail$ axis and evaluate the full-covariance $\chi^{2}$ of the unfolded
        data against each generator's $d\sigma/d\Eavail$: all four are incompatible
        with the data at $>21\sigma$ over the seven $\Eavail$ bins and $>15\sigma$ in
        the DIS tail ($\Eavail\ge\SI{0.8}{GeV}$) alone --- GENIE~CV $22.4\sigma$,
        GENIE\,$+$\,MEC $24.9\sigma$, NuWro $21.9\sigma$, GiBUU $21.2\sigma$. GiBUU
        spreads its deficit across the \emph{entire} DIS tail (data/generator
        $1.59$ at $\Eavail\in[0.8,1.5)$, $1.36$ at $[1.5,3.0)$) rather than piling
        it in the deep-DIS catch bin like the GENIE variants, an independent and
        qualitatively distinct confirmation. Resolving the excess in $W$ confirms
        and sharpens this picture: we build the full $(\Eavail,W)$ covariance
        (42 bins; \texttt{C\_syst}$+$\texttt{C\_stat}$+$\texttt{C\_lateral}, median
        \SI{14.9}{\percent} per bin) by a frozen-reweighter block sum on the 5D
        \texttt{\_universes\_full} sample with a $W$-resolved lateral block
        (\S\ref{sec:eavailw-cov}), validated by reproducing the frozen 5D
        marginals ($\max|\text{ratio}-1|=0.007$). Over the full $(\Eavail,W)$ plane
        the four generators are incompatible with the data at GENIE~CV $19.3\sigma$
        ($\chi^{2}/\mathrm{ndf}=515.1/42$), GENIE\,$+$\,MEC $19.0\sigma$ ($502.0/42$),
        NuWro $35.9\sigma$ ($1479.2/42$), and GiBUU $>40\sigma$ ($2466.7/42$). The
        deficit is \emph{localised in the high-$W$ DIS corner}: restricting to the
        twelve bins with $\Eavail\ge\SI{0.4}{GeV}$ and $W\ge\SI{1.8}{GeV}$, all four
        miss the data at $8.9$--$22.4\sigma$ --- GENIE~CV $8.9\sigma$ ($114.3/12$),
        GENIE\,$+$\,MEC $9.2\sigma$ ($121.1/12$), NuWro $15.6\sigma$ ($290.3/12$), and
        GiBUU $22.4\sigma$ ($556.6/12$, again the most deficient). The excess is thus
        a genuinely DIS-region ($W\gtrsim\SI{1.8}{GeV}$) feature, not a low-$W$
        resonance artefact. \emph{The low-recoil cross-check is done:} the
        bin-identical full-covariance comparison to
        Ref.~\cite{MINERvA:2022incl} on the maximal common $(\Eavail,q_{3})$
        grid gives $\chi^{2}/\mathrm{ndf}=1.68/2$, $p=0.43$ (\S\ref{sec:4d}).
\end{enumerate}
