\documentclass[11pt]{article}
\usepackage{authblk}
\input{preamble}
\NOTEtrue
\input{values}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,      
    urlcolor=blue,
    citecolor=blue,
}

\usepackage{xcolor}
\newcommand{\bpn}[1]{\textcolor{blue}{\textbf{[BPN: #1]}}}
\newcommand{\jrb}[1]{\textcolor{violet}{\textbf{[JRB: #1]}}}


%\title{Unbinned OmniFold measurement of the MINERvA ME FHC\\
%CC-inclusive double-differential cross section\\
%$d^{2}\sigma/(d\pT\,d\pPar)$ and its higher-dimensional extensions\\[4pt]
%\large MINERvA analysis note --- DRAFT}

\title{\textbf{Analysis Note:}\\
Unbinned measurement of the charged-current inclusive cross section in vN in collisions with public MINERvA data}

\author[1]{Joseph Bailey\thanks{\url{josephrb@stanford.edu}}}
\author[2]{Benjamin Nachman\thanks{\url{nachman@stanford.edu}}}
\affil[1]{Department of Physics, Stanford University}
\affil[2]{Department of Particle Physics and Astrophysics, Stanford University and SLAC National Laboratory}


\date{\today~~DRAFT v1.0\\[4pt]\small Code and file references: \repolink}

\begin{document}
\maketitle

\begin{abstract}

We perform unbinned, multi-dimensional cross-section measurements of medium-energy charged-current neutrino-nucleus events measured by the MINERvA experiment.  The machine learning-based method OmniFold is used to enable these measurements.  We first reproduce a published double-differential cross section measurement in muon transverse and longitudinal momentum, recovering both the spectrum and its uncertainty budget with a fully independent method. We then perform the first simultaneous three-, four-, and five-observable measurements of MINERvA data, adding available energy ($\Eavail$), three-momentum transfer ($q_{3}$), and hadronic invariant mass ($W$); each reproduces its lower-dimensional predecessor under marginalization. These higher-dimension results recover, through an independent method, the established MINERvA deficit of low-momentum-transfer multinucleon interactions, and localize a residual high-$\Eavail$ excess to the deep-inelastic region, where no current generator reproduces the data. A recoil-only point-cloud transformer provides a low-level-representation cross-check, while a scalar study explores an extended fiducial phase space obtained by removing truth-level muon kinematic cuts. The latter separates an acceptance-supported extension from a prior-dominated extrapolation region and does not claim that the entire enlarged truth domain is uniformly measured. The note records the validated central results, the corrected 5D GBDT uncertainty campaign, the completed recoil-PET cross-check budget, and the validation suite; a genuine full-event PET estimator and corrected extended-fiducial uncertainties remain in progress.

\end{abstract}

\input{sec_execsummary}
\clearpage

\tableofcontents

\input{sec_intro}
\input{sec_experiment}
\input{sec_method}
\input{sec_systematics}
\input{sec_results}
\input{sec_validation}
\input{sec_3d}
\input{sec_eavailw}
\input{sec_pet}
\input{sec_fps}
\input{sec_summary}

\section*{Acknowledgments}
This
analysis was carried out on the Perlmutter system at the National Energy
Research Scientific Computing Center (NERSC), a U.S.\ Department of Energy
Office of Science User Facility, under project m3246. It builds on the MINERvA
Analysis Toolkit (MAT) and the MINERvA-101 cross-section framework, and I thank
their authors. I am grateful to the MINERvA Collaboration for the public release
of the medium-energy CC-inclusive data product on which the reproduction rests.

We thank Callum Wilkinson for connecting us with this dataset and for many helpful discussions.

Large language model assistants (Anthropic's Claude and OpenAI's ChatGPT) were used to support brainstorming and methodological discussion, code development, analysis scripting, and manuscript preparation. All scientific ideas, results, and claims were critically evaluated and verified by the authors, who take full responsibility for the content.

\appendix
\input{app_statmethods}
\input{app_negweight}
\input{app_landscape}
\input{app_codebase}

\printbibliography

\end{document}
