\section*{Executive summary}
\addcontentsline{toc}{section}{Executive summary}
\label{sec:execsummary}

This note documents the application of \emph{unbinned machine-learning unfolding} (OmniFold~\cite{Andreassen:2019cjw, Andreassen:2021zzk}) to MINERvA's medium-energy neutrino data --- first as a faithful reproduction of a published flagship measurement, then as its extension to simultaneous three-, four- and five-observable scalar cross sections, a recoil-point-cloud representation cross-check, and a preliminary extended-fiducial study with relaxed truth-muon cuts. None has a binned counterpart at MINERvA. We reserve \emph{full-event} for an estimator whose classifier inputs contain the declared muon and recoil information together; the existing PET products do not meet that definition. The executive summary outlines the whole story; everything afterwards is the supporting technical record.

Traditional unfolding corrects detector effects with a migration matrix between histogram bins, so each added observable multiplies the number of bins and the method drowns in dimensions. OmniFold instead trains a classifier to learn, event by event, a weight that morphs the simulation into the data, and pulls that weight back to the true (pre-detector) level through the simulation's truth--reco pairing. No binning is imposed until the very end, so adding an observable is just adding an input column. This analysis uses gradient-boosted decision trees for the fixed-dimension measurements and
validates the choice extensively (closure tests, coverage, calibrated-classifier checks, and neural-network cross-checks).

\paragraph{Uncertainty-remediation status.}
The validated 2D reproduction, higher-dimensional central values, closure
tests, marginalization anchors, and shape-localization statements remain
current. The corrected 5D GBDT covariance is now adopted. The present PET
campaign has completed a five-component uncertainty assembly for the existing
recoil-only estimator; none of its components transfers automatically to the
full-event estimator now being implemented. Corrected 4D and extended-fiducial
covariances and covariance-based generator significances remain withheld.

\paragraph{Status of the results in this draft.}
\begin{center}
\small
\begin{tabular}{@{}p{0.23\textwidth}p{0.69\textwidth}@{}}
\toprule
Category & Contents \\
\midrule
Validated/current & Final 2D reproduction; scalar 3D/4D/5D central values and marginalization anchors; corrected 5D GBDT covariance; qualitative localization of the high-$\Eavail$, high-$W$ excess. \\
\addlinespace
Preliminary or representation cross-check & Scalar extended-fiducial central value, anchor and validation controls while final UQ is pending; all existing recoil-only PET and PET-on-extended-grid products, including their closures and uncertainty diagnostics. \\
\addlinespace
Planned/in progress & Full-event PET over the declared extended fiducial phase space, omitted-muon stress closure, fresh nominal and UQ, selection-complete laterals, corrected extended-fiducial covariance, and covariance-dependent significances. \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Results overview:}
\begin{itemize}[itemsep=2pt]
  \item \textbf{Reproduction.} 
  Our unbinned measurement of the muon transverse and longitudinal momentum,  $d^{2}\sigma/(d\pT\,d\pPar)$, matches the published measurement~\cite{MINERvA:2021owq}: total cross-section ratio $\ratioTot$, \SI{\binsTen}{\percent} of bins within \SI{10}{\percent}, per-bin pulls sub-$1\sigma$, and a fully independent uncertainty budget whose median per-bin uncertainty (\SI{\uqMedian}{\percent}) matches the published \SI{\uqPaper}{\percent} (\S\ref{sec:results}, \S\ref{sec:systematics}, Fig.~\ref{fig:xsec}).
  
  \item \textbf{First 3D/4D/5D MINERvA measurements.} 
  Adding available energy ($\Eavail$), momentum transfer ($q_{3}$), and hadronic mass ($W$) produces the first simultaneous three-, four- and five-observable unfolds of MINERvA data; every result reproduces its lower-dimensional predecessor when marginalized --- a built-in cross-check (\S\ref{sec:3d}).
  
  \item \textbf{Known physics recovered independently.} 
  All four truth-level generators (GENIE~\cite{Andreopoulos:2009rq}, MINERvA Tune~\cite{MINERvA:2016lowrecoil}, NuWro~\cite{Golan:2012wx}, GiBUU~\cite{Buss:2011mx}) under-predict the data at low $\Eavail$; pion final-state-interaction dials move the spectrum by $<\SI{1}{\percent}$ while adding Valencia 2p2h~\cite{Nieves:2011pp,Gran:2013kda} fills the quasielastic--$\Delta$ dip --- the established MINERvA low-recoil/2p2h deficit~\cite{MINERvA:2016lowrecoil,MINERvA:2022incl}, re-found by an independent method (\S\ref{sec:3d-2p2h}, Fig.~\ref{fig:3dmodels}).
  
  \item \textbf{A new localization of the remaining excess.} 
  The high-$\Eavail$ excess that 2p2h cannot explain is localized in the $(\Eavail, W)$ plane: its positive data-minus-generator difference concentrates in the deep-inelastic corner ($W\ge\SI{1.8}{GeV}$), and all four generators underpredict that region. Exact significances are withheld while the corrected covariance is regenerated (\S\ref{sec:eavailw}, Fig.~\ref{fig:eavailWband}).
  
  \item \textbf{A methodological finding.} 
  The standard covariance adds matrices built for one systematic source at a time, which omits cross-source response terms. In MINERvA's \texttt{GetTotalErrorMatrix}, each named source contributes its own covariance, built with all other sources held nominal; the sum retains within-source bin--bin correlations but contains no cross-source terms. The corrected unified-throw test perturbs all reweightable sources simultaneously using actual asymmetric endpoints, one fixed estimator seed, and mean-centered throws; it reports the joint mean shift separately. In 5D the corrected block sum has median uncertainty \SI{\gbdtFiveBlockMedian}{\percent}, while the adopted mean-centered covariance has $\sqrt{\mathrm{Tr}\,C}=\SI{\gbdtFiveAdoptTrace}{cm^2/nucleon}$; the CV-centered \SI{\gbdtFiveCVTrace}{cm^2/nucleon} variant is retained as a conservative alternative (\S\ref{sec:uq-combine}).
  
  \item \textbf{Recoil-point-cloud representation cross-check.}
  A point-cloud transformer operates on non-muon reconstructed recoil clusters at step~1 and truth final-state hadrons at step~2. Muon kinematics enter selection and downstream binning but not either classifier, so the learned weight is not a weight over the complete joint muon--hadron event. The absolutely normalized central value, ordinary closure and completed corrected 5D uncertainty assembly remain useful diagnostics of this recoil-only estimator; ordinary closure validates the extraction machinery but cannot test omitted muon dependence at fixed recoil. A distinct full-event interface and stress-closure gate are now being implemented, after which the nominal and UQ must be repeated (\S\ref{sec:pet}, \S\ref{sec:pet-projection}).

  \item \textbf{A preliminary extended-fiducial study.}
  Re-running the scalar 2D $(\pT,\pPar)$ unfold with the truth-muon kinematic cuts removed enlarges the signal definition while retaining the tracker fiducial interaction volume and the unchanged MINOS-based reconstructed selection. The preliminary central value, restricted-region anchor and validation controls remain current. About \SI{6}{\percent} of the rate lies in a newly acceptance-supported extension (tier~1), while about \SI{28}{\percent} lies in low- or zero-efficiency cells and is reported separately as prior-dominated extrapolation (tier~2). A more expressive classifier cannot create detector information in those cells, and corrected final UQ is still pending. The existing PET run on this population remains a recoil-only feasibility study, not the full-event result (\S\ref{sec:fps}).
 
  
\end{itemize}

This note is organized as follows.  Sections~\ref{sec:experiment}--\ref{sec:method} define the samples and the method; \S\ref{sec:systematics} and App.~\ref{app:statmethods} are the uncertainty machinery (the appendix documents the analysis-specific procedures --- variance sources, coverage and closure toys, the covariance construction --- for reproduction); \S\ref{sec:results} is the 2D reproduction against the published measurement; \S\ref{sec:validation} is the validation suite; the physics is in \S\ref{sec:3d}--\ref{sec:eavailw}; \S\ref{sec:pet} is the recoil-point-cloud cross-check; \S\ref{sec:fps} is the extended-fiducial study and \S\ref{sec:summary} the outlook.
